Pre-/Post-Trikafta in CFRD

### 2022-02-02 Christine Chan/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "B:/Projects"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/BDC/Projects"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}

```

>>>>

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lmerTest)
library(broom.mixed)
library(performance)
library(lubridate)
library(plotly)
library(knitr)
knitr::opts_chunk$set(echo = FALSE,fig.height = 8,fig.width = 10)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF"
}
knitr::opts_knit$set(root.dir = home_dir)
```

```{r include=FALSE}
# Import cleaned data from Casey
df = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
df$X = NULL
# Fill in MRN and CFF ID for everyone
df = df %>% 
  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
# Assume that blanks for pancreatic and G tube indicate "no"
df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
# Convert dates
date_vars = c("DOB","Start","Date")
df[,date_vars] = lapply(df[,date_vars], function(d){
  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
})
# Time from start and filtering
df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
# Add UM (they did not provide dates, so need Time column)
um_a1cs = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
um_bmi = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
um_demo = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
um_ogtt = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
um_pfts = read.csv("./Christine Chan/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
um = full_join(um_a1cs,um_bmi)
um = full_join(um,um_ogtt)
um = full_join(um,um_pfts)
um = left_join(um,um_demo)
um = um %>% filter(!is.na(CFRD.Status))
df = full_join(df,um)
# Filter and fill constants
df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
```

```{r}
data = df
outcome = "HbA1c"
# Piecewise model function
piecewise = function(outcome,data = df){
  # Add spline column
  data$Change = ifelse(data$Time < 0,0,1)
  data$Change = data$Change * data$Time
  # Model
  f = as.formula(paste0(outcome,"~Time + Change + (1|CFF.ID)"))
  mod = lmer(f,data = data)
  res = tidy(mod,"fixed")
  res$effect = NULL
  # Coordinates for model plotting
  x1_start = -365
  y1_start = res$estimate[1] + res$estimate[2]*x1_start
  x1_end = 0
  y1_end = res$estimate[1] 
  x2_start = 0
  y2_start = res$estimate[1] 
  x2_end = 365
  y2_end = res$estimate[1] + (res$estimate[2]+res$estimate[3])*x2_end
  # Plot
  p = ggplot(data,aes_string(x = "Time",y = outcome,group = "CFF.ID",color = "Site")) +
    geom_point(alpha = 0.5) + geom_line(alpha = 0.2) +
    annotate("segment",x=x1_start,y=y1_start,xend=x1_end,yend=y1_end,color = "blue",size = 1) +
    annotate("segment",x=x2_start,y=y2_start,xend=x2_end,yend=y2_end,color = "blue",size = 1) +
    xlab("Days from Trikafta Start") + theme_bw()
  # Interactive plot and model results
  return(list(plot = ggplotly(p),table = res))
}
```

# Methods

# HbA1c

```{r results='asis'}
a1c = piecewise(outcome = "HbA1c",df)
ggplotly(a1c$plot)
kable(a1c$table,digits = 3)
```
